---
title: "Getting Started with CDSim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with CDSim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  fig.height = 4,
  comment = "#>"
)
library(CDSim)
```
The CDSim package provides an easy workflow for simulating climate data such as temperature and rainfall across multiple synthetic weather stations. It is useful for testing models, teaching climate analysis, generating demo data, and creating datasets with controlled variability.

This vignette demonstrates:

 - how to create synthetic weather stations

 - how to generate multi-year climate time series

 - how to export results to CSV and NetCDF

 - how to perform a quick visualization

<h2 style="font-size:160%; font-weight:bold;"> Creating Weather Stations </h2>

The stations can be created either by:

  - loading from a CSV file,
  
  - accepting an existing data frame,
  
  - or auto-generating synthetic stations in a bounding box.
  
```{r}
library(CDSim)
```
We begin by generating a set of synthetic weather stations. The seed ensures
reproducibility.
```{r}
stations <- create_stations(n = 3, seed = 123)
stations
```
Each station typically contains:

  1. station name

  2. longitude

  3. latitude
  
<h2 style="font-size:160%; font-weight:bold;"> Simulating Climate Data </h2>

Once the stations are created, we can generate daily or monthly climate time
series using built-in stochastic models.
```{r}
sim <- simulate_climate_series(stations, start_year = 2019, end_year = 2024)
head(sim)
```
A typical simulated record includes:

  - date

  - min and max temperature

  - rainfall

  - station metadata
  
<h2 style="font-size:160%; font-weight:bold;"> Exporting Data to CSV and NetCDF </h2>

CDSim includes convenient exporters such as CSV and NetCDF for storing climate data. This makes it easier for packages such as ncdf4, terra, or stars to read the outputs.
```{r}
write_station_csv(sim, file = "climate_data.csv")
```
```{r}
write_station_netcdf(sim, out_nc = "climate_data.nc")
```
<h2 style="font-size:160%; font-weight:bold;"> Quick Visualization </h2>

To demonstrate a quick plot, here’s the maximum temperature series of the first station.
```{r}
plot_station_timeseries(sim,'Station_1', var = "Avg.Tx")
```
